---
title: "Lab 3"
author: "Christian Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 3 Tasks

## Task 1

```{r}
getwd()
setwd("C:/Users/cmill/OneDrive/Desktop/Applied Stats/Labs/Lab 3")

```

## Task 2

```{r}
spruce <- read.csv("SPRUCE.csv")

head(spruce)

file.create("mylab3.R")
```
## Task 3
```{r}

heightplot <- plot(Height ~ BHDiameter, bg = "blue", pch = 21, cex = 1.2, ylim = c(0, 1.1*max(Height)), xlim = c(0, 1.1*max(BHDiameter)), xlab = "BHDiameter", ylab = "Height", main = "Height vs. Breast Height Diameter", data = spruce)
```

There is not a straight line relationship here as there is too much variability in the data, though the data points in the middle appear to be more linear.

```{r}
library(s20x)

lo <- layout(matrix(1:3, nrow = 1, ncol = 3, byrow = TRUE))
tsone <- trendscatter(Height ~ BHDiameter, f = 0.5, data = spruce)
tstwo <- trendscatter(Height ~ BHDiameter, f = 0.6, data = spruce)
tsthree <- trendscatter(Height ~ BHDiameter, f = 0.7, data = spruce)

spruce.lm <- with(spruce, lm(Height ~ BHDiameter))

lay <- layout(1)

with(spruce,
     plot(Height ~ BHDiameter, main = "Tree Height vs BHDiameter",
          xlab = "Breast Height Diameter (cm)",
          ylab = "Height of Tree (m)",
          bg = "blue", pch = 21, cex = 1.2,
          ylim = c(0, 1.1 * max(Height)), xlim = c(0, 1.1 * max(BHDiameter)))
)
abline(spruce.lm)
```

The straight line is not appropriate in this case and the trendscatter plot shows a curvature for each of the f-values.

## Task 4
```{r}
layout1 <- layout(matrix( 1:4, nr = 2, nc = 2, byrow = TRUE))

layout.show(4)

mycex <- 1.0

with(spruce,
     plot(Height ~ BHDiameter, main = "Tree Height vs BHDiameter",
          xlab = "Breast Height Diameter (cm)",
          ylab = "Height of Tree (m)",
          bg = "blue", pch = 21, cex = mycex,
          ylim = c(0, 1.1 * max(Height)), xlim = c(0, 1.1 * max(BHDiameter)))
)
abline(spruce.lm)

yhat <- with(spruce, predict(spruce.lm, data.frame(BHDiameter)))

with(spruce,
     plot(Height ~ BHDiameter, main = "RSS",
          xlab = "Breast Height Diameter (cm)",
          ylab = "Height of Tree (m)",
          bg = "blue", pch = 21, cex = mycex,
          ylim = c(0, 1.1*max(Height)), 
          xlim = c(0, 1.1*max(BHDiameter)))
     )
with(spruce, {
  segments(BHDiameter, Height, BHDiameter, yhat)
})
abline(spruce.lm)

with(spruce,
     plot(Height ~ BHDiameter, main = "MSS",
          xlab = "Breast Height Diameter (cm)",
          ylab = "Height of Tree (m)",
          bg = "blue", pch = 21, cex = mycex,
          ylim = c(0, 1.1*max(Height)),
          xlim = c(0, 1.1*max(BHDiameter)))
     )
with(spruce, abline(h = mean(Height)))

with(spruce, segments(BHDiameter, mean(Height), BHDiameter, yhat, col = "red"))
abline(spruce.lm)

with(spruce,
     plot(Height ~ BHDiameter, main = "TSS Plot by Christian Miller",
          xlab = "Breast Height Diameter (cm)",
          ylab = "Height of Tree (m)",
          bg = "blue", pch = 21, cex = mycex,
          ylim = c(0, 1.1*max(Height)),
          xlim = c(0, 1.1*max(BHDiameter)))
)
with(spruce, abline(h = mean(Height)))
with(spruce, segments(BHDiameter, Height, BHDiameter, mean(Height), col = "green"))

```

### Total Sum of Squares
```{r}
TSS <- with(spruce, sum((Height - mean(Height))^2))
TSS
```
The TSS is 278.9475

### Model Sum of Squares
```{r}
MSS <- with(spruce, sum((yhat - mean(Height))^2))
MSS
```
The MSS is 183.2447

### Residual Sum of Squares
```{r}
RSS <- with(spruce, sum((Height - yhat)^2))
RSS
```
The RSS is 95.70281

### R-Squared
```{r}
R2 <- MSS/TSS
R2
```
The R-squared is equal to 0.6569146, meaning that roughly 65.7% of the variability in the output can be explained by the model.

```{r}
RSS + MSS
TSS
```
Yes, the RSS + MSS = TSS (equating to 278.9475)

## Task 5
```{r}
summary(spruce.lm)
```
The value of the slope is 0.48147 and the intercept is 9.14684.

The equation of the fitted line is: E(Height) = 9.14684 + 0.48147*BHDiameter

```{r}
predict(spruce.lm, data.frame(BHDiameter = c(15, 18, 20)))
```
The model predicts that the height of spruce will be 16.36895 meters when the diameter is 15 cm.
The model predicts that the height of spruce will be 17.81338 meters when the diameter is 18 cm.
The model predicts that the height of spruce will be 18.77632 meters when the diameter is 20 cm.

## Task 6
```{r}
library(ggplot2)

g <- ggplot(spruce, aes(x = BHDiameter, y = Height, color = BHDiameter))
g <- g + geom_point() + 
  geom_line() +
  geom_smooth(method = "lm")
g = g + ggtitle("Tree Height vs Breast Height Diameter - Plot by Christian Miller")
g

```


## Task 7


