---
title: "Lab 5"
author: "Christian Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 4 Tasks

## Task 1
```{r}
setwd("C:/Users/cmill/OneDrive/Desktop/Applied Stats/Labs/Lab 5")
getwd()
```

## Task 2
```{r}
file.create("mylab5.r")
```

### mybin
```{r}
mybin=function(iter=1000,n=10, p=0.5){ 
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  succ=c()
  for( i in 1:iter){
    sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
    succ[i]=sum(sam.mat[,i])
}
  succ.tab=table(factor(succ,levels=0:n))
  barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation",   xlab="Number of successes")
    succ.tab/iter
}
```

```{r}
bin100 <- mybin(iter = 100, n = 10, p = 0.7)
bin200 <- mybin(iter = 200, n = 10, p = 0.7)
bin500 <- mybin(iter = 500, n = 10, p = 0.7)
bin1000 <- mybin(iter = 1000, n = 10, p = 0.7)
bin10000 <- mybin(iter = 10000, n = 10, p = 0.7)
```

#### 10,000 Iterations 

```{r}
bin10000

dbinom(0:10, size = 10, prob = 0.7)
```
 
Comparing the values observed from the simulation after iterating the experiment 10,000 times with the theoretical probabilities, we see that they are approximately consistent with the theoretical values.
 
## Task 3
 
### Marble Experiment
 
#### Without Replacement
 
```{r}
marbles <- c(rep(1, 12), rep(0, 8))

sample_marbles <- sample(marbles, size = 5, replace = FALSE)

print(sample_marbles)
```
 
#### With Replacement
 
```{r}
sample_marbles2 <- sample(marbles, size = 5, replace = TRUE)

print(sample_marbles2)
```
 
#### Hypergeometric Distribution

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
succ=c()
for( i in 1:iter){
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
succ[i]=sum(sam.mat[,i])
}
succ.tab=table(factor(succ,levels=0:n))
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}

hyper100 <- myhyper(iter=100,n=5, N=20,r=12)

hyper200 <- myhyper(iter=200,n=5, N=20,r=12)

hyper500 <- myhyper(iter=500,n=5, N=20,r=12)

hyper1000 <- myhyper(iter=1000,n=5, N=20,r=12)

hyper10000 <- myhyper(iter=10000,n=5, N=20,r=12)
```

##### 10,000 Iterations

```{r}
hyper10000
?dhyper
dhyper(x=0:5, m=12, n=8, k=5)
```

Comparing the tables, we can verify that the simulation approximately reflects the theoretical values.

## Task 4

### mysample()

The `mysample` function iteratively samples from numbers 1 to 10, visualizing each sample's proportions in a barplot to illustrate sampling variability and randomness in sampling from a uniform distribution. It repeats this visualization across specified iterations, highlighting how sample outcomes can vary even under consistent conditions.

```{r}
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){
s=sample(1:10,n,replace=TRUE)
sf=factor(s,levels=1:10)
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)
Sys.sleep(time)
}
}

```

Running `mysample(n=1000, iter=30, time = 1)` returned 30 bar plots, each showing approximately uniform proportions of each number from 1 to 10. This demonstrates how large samples reduce sampling variability, leading to consistent and predictable outcomes.

Here is an example of one of the plots rendered from the `mysample()` function:

```{r}
mysample(n=1000, iter=1, time = 1)
```

## Task 5

### 8 Choose 4

```{r}
choose(8, 4)
```

### Other Calculations

$$
P(Y>4),\ Y \sim \text{Pois}(\lambda=2)
$$
```{r}
1 - ppois(4, lambda = 2)
```


$$
P(Y=10),\ Y \sim \text{NegBin}(p=0.4, r=3)
$$
```{r}
mynbin=function(y,r,p){
choose(y-1,r-1)*p^r*(1-p)^(y-r)
}
mynbin(10,3,0.4)
```

### 
$$
P(Y \leq 8),\ Y \sim \text{Bin}(n=15, p=0.4)
$$
```{r}
pbinom(8, size = 15, prob = 0.4)
```

## Task 6

### mybin() function

```{r}
MATH4753CSM24::mybin(iter = 100, n = 10, p = 0.5)
```

