---
title: "Lab 7"
author: "Christian Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 7 Tasks

## Task 1

```{r}
setwd("C:/Users/cmill/OneDrive/Desktop/Applied Stats/Labs/Lab 7")
getwd()
```

## Task 2: The Chi-square Statistic

```{r}
mychisim<-function(n1=10,sigma1=3,mean1=5,iter=1000,ymax=0.1,x=20, y=0.1){    # adjust ymax to make graph fit
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1

data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1

ssq1=apply(data1.mat,2,var) # ssq1 is s squared

w=(n1-1)*ssq1/sigma1^2      #chi-sq stat

hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",chi^2)),
xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dchisq(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(chi^2==frac((n[1]-1)*s^2,sigma^2)) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq")) # some output to use if needed
}
```

### Sampling from a Simulated Population

```{r}
chisq=mychisim(iter=1000, mean1= 10, sigma1 = 4, ymax=0.15, x = 20, y = 0.1)
```

```{r}
chisq=mychisim(n1 = 20, iter=1000, mean1= 10, sigma1 = 4, ymax=0.1, x = 30, y = 0.08)
```


```{r}
chisq=mychisim(n1 = 100, iter=1000, mean1= 10, sigma1 = 4, ymax = 0.04, x = 120, y = 0.03)
```

```{r}
chisq=mychisim(n1 = 200, iter=1000, mean1= 10, sigma1 = 4, ymax=0.05, x = 230, y = 0.035)
```


```{r}
chisq=mychisim(n1 = 10, iter=1500, mean1= 20, sigma1 = 10, ymax=0.10)

hist(chisq$w, col = rainbow(15), xlab = expression(paste(Chi^2, "Value", sep = " "), las=1))
```


## Task 3: The Student's T Statistic

```{r}
myTsim<-function(n1=10,sigma1=3,mean1=5,iter=1000,ymax=0.1,x=2,y=0.3,...){    # adjust ymax to make graph fit
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1

data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1

sd1=apply(data1.mat,2,sd) # sd
ybar=apply(data1.mat,2,mean)  # mean

w=(ybar-mean1)/(sd1/sqrt(n1))      #T stat

hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",T," iterations= ",iter)),
xlab=expression(paste(T, "Statistic",sep=" ")), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dt(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(T==frac((bar(y)-mu),s/sqrt(n1))) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="T")) # some output to use if needed
}
```

### Sampling from a Simulated Population

```{r}
myTsim(n1 = 10, iter=1000, mean1 = 10, sigma1 = 4, ymax=0.45)
```

```{r}
myTsim(n1 = 20, iter=1000, mean1 = 10, sigma1 = 4, ymax=0.45)
```


```{r}
myTsim(n1 = 100, iter=1000, mean1 = 10, sigma1 = 4, ymax=0.45, x = 1.5, y = 0.35)
```


```{r}
myTsim(n1 = 200, iter=1000, mean1 = 10, sigma1 = 4, ymax=0.5, y = 0.4)
```


```{r}
T = myTsim(n1 = 10, iter=1500, mean1 = 20, sigma1 = 10, ymax=0.5, y = 0.4)
hist(T$w, col = rainbow(15), xlab = expression(paste(T, "-Value", sep = " "), las=1))
```


## Task 4: Chi-square from Two Population Sampling

```{r}
mychisim2<-function(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000,ymax=0.07,x=40,y=0.04,...){    # adjust ymax to make graph fit
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
ssq1=apply(data1.mat,2,var) # ssq1 is s squared
ssq2=apply(data2.mat,2,var)
spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
w=(n1+n2-2)*spsq/(sigma1^2)#sigma1=sigma2,  Chi square stat
hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",chi^2)),
xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dchisq(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(chi^2==frac((n[1]+n[2]-2)*S[p]^2,sigma^2)) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq")) # some output to use if needed
}
```

### Sampling from Two Simulated Populations

```{r}
mychisim2(n1 = 10, n2 = 10, mean1 = 5, mean2 = 10, sigma1 = 4, sigma2 = 4, iter=1000, x = 30, y = 0.05)
```


```{r}
mychisim2(n1 = 20, n2 = 10, mean1 = 3, mean2 = 5, sigma1 = 10, sigma2 = 10, iter=1000, x = 45, y = 0.05)
```

```{r}
mychisim2(n1 = 50, n2 = 50, mean1 = 5, mean2 = 10, sigma1 = 4, sigma2 = 4, iter=10000, ymax = 0.04, x = 130, y = 0.035)
```

```{r}
mychisim2(n1 = 80, n2 = 50, mean1 = 3, mean2 = 5, sigma1 = 10, sigma2 = 10, iter=10000, ymax = 0.04, x = 160, y = 0.035)
```

```{r}
chisq2 = mychisim2(iter = 10000)
hist(chisq2$w, col = rainbow(15), xlab = expression(paste(Chi^2, "Value", sep = " "), las=1))
```



## Task 5: Student's T From Two Population Sampling

### Altering the Legend Option in myTsim2() 

```
legend(locator(1),c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)
```


### The Student's T Statistic and Notation

$$
T = \frac{(\bar{Y}_1 - \bar{Y}_2) - (\mu_1 - \mu_2)}{S_{p}\sqrt{\frac{1}{n_1} + \frac{1}{n_2}}}
$$

where

$$
S^2_{p} = \frac{(n_1 - 1)S^2_1 + (n_2 - 1)S^2_2}{n_1 + n_2 - 2}
$$

where:

- \( \bar{Y}_1 \) is the sample mean of the first group.
- \( \bar{Y}_2 \) is the sample mean of the second group.
- \( \mu_1 \) is the population mean of the first group under the null hypothesis.
- \( \mu_2 \) is the population mean of the second group under the null hypothesis.
- \( S_{p} \) is the pooled standard deviation of the two groups.
- \( n_1 \) is the sample size of the first group.
- \( n_2 \) is the sample size of the second group.
- \( S_p^2 \) is the pooled sample variance.

Within the ```myTsim2()``` function, ```w=((ybar1-ybar2)-(mean1-mean2))/sqrt(spsq*(1/n1+1/n2))``` calculates the T-statistic.

```{r}
myTsim2<-function(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000,ymax=0.5,x=2,y=0.4,...){
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
ssq1=apply(data1.mat,2,var) # ssq1 is s squared
ybar1= apply(data1.mat,2,mean)
ssq2=apply(data2.mat,2,var)
ybar2=apply(data2.mat,2,mean)
spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
w=((ybar1-ybar2)-(mean1-mean2))/sqrt(spsq*(1/n1+1/n2))#sigma1=sigma2,  Chi square stat
hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",T)),
xlab=paste(" T Statistic",sep=""), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dt(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(T==frac((bar(Y)[1]-bar(Y)[2])-(mu[1]-mu[2]),S[p]*sqrt(frac(1,n[1])+frac(1,n[2])))) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
invisible(list(w=w,summary=summary(w),sdw=sd(w),fun="T")) # some output to use if needed
}
```


### Sampling from Two Simulated Populations

```{r}
myTsim2(n1 = 10, n2 = 10, mean1 = 5, mean2 = 10, sigma1 = 4, sigma2 = 4, iter = 1000)
```


```{r}
myTsim2(n1 = 20, n2 = 10, mean1 = 3, mean2 = 5, sigma1 = 10, sigma2 = 10, iter = 1000, x = 1.5)
```


```{r}
myTsim2(n1 = 50, n2 = 50, mean1 = 5, mean2 = 10, sigma1 = 4, sigma2 = 4, iter = 10000)
```


```{r}
myTsim2(n1 = 80, n2 = 50, mean1 = 3, mean2 = 5, sigma1 = 10, sigma2 = 10, iter = 10000)
```


```{r}
T2 <- myTsim2(iter = 10000)
hist(T2$w, col = rainbow(15), xlab = expression(paste(T, "-Value", sep = " "), las=1))
```


## Task 6: The F Statistic from Two Population Sampling

### The F Statistic and Notation

The F statistic is used to compare the variances of two populations and is calculated as the ratio of two sample variances. It is defined as:

$$
F = \frac{\left(\frac{S_1^2}{\sigma_1^2}\right)}{\left(\frac{S_2^2}{\sigma_2^2}\right)}
$$

where:

- \( S_1^2 \) is the variance of the first sample.
- \( S_2^2 \) is the variance of the second sample.
- \( \sigma_1^2 \) is the variance of the first population under the null hypothesis.
- \( \sigma_2^2 \) is the variance of the second population under the null hypothesis.
- \( \nu_1 = (n_1 - 1) \) is the numerator degrees of freedom, which corresponds to the first sample size minus one.
- \( \nu_2 = (n_2 - 1) \) is the denominator degrees of freedom, which corresponds to the second sample size minus one.

The F statistic is based on the following assumptions:

- The samples are drawn independently from their respective populations.
- The samples are random, meaning every member of the population has an equal chance of being included in the sample.
- The samples come from normally distributed populations.
- The populations from which the samples are drawn have the parameters \( \mu_1, \sigma_1^2 \) for the first population and \( \mu_2, \sigma_2^2 \) for the second population.

Within the ```myFsim2()``` function, ```w=ssq1*sigma2^2/(ssq2*sigma1^2)``` calculates the F-statistic.

```{r}
myFsim2<-function(n1=10,n2=14,sigma1=3,sigma2=2,mean1=5,mean2=10,iter=1000,ymax=0.9,x=6,y=0.5,...){
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
ssq1=apply(data1.mat,2,var) # ssq1 is s squared
ssq2=apply(data2.mat,2,var)
#spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
w=ssq1*sigma2^2/(ssq2*sigma1^2) #
hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",F)),
xlab=paste("F Statistic",sep=""), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(df(x,n1-1,n2-1),xlim=c(0,6),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(F==frac(s[1]^2,s[2]^2)*frac(sigma[2]^2,sigma[1]^2)) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="F")) # some output to use if needed
}
```


### Sampling from Two Simulated Populations

```{r}
myFsim2(n1=10,n2=14,sigma1=3,sigma2=2,mean1=10,mean2=20,iter=1000,ymax=0.9, x = 4)
```


```{r}
myFsim2(n1=14,n2=10,sigma1=2,sigma2=3,mean1=10,mean2=5,iter=1000,ymax=0.9, x = 6, y = 0.6)
```


```{r}
myFsim2(n1=50,n2=10,sigma1=5,sigma2=10,mean1=10,mean2=10,iter=1000,ymax=1)
```


```{r}
myFsim2(n1=5,n2=5,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000,ymax=0.6)
```


```{r}
myFsim2(n1=10,n2=14,sigma1=3,sigma2=2,mean1=5,mean2=10, iter=1000, ymax=0.9, x = 4)
```


## Task 7: Adding Data and Documentation to Package

```{r}
library(MATH4753CSM24)
data("fire")
knitr::kable(head(fire))
?fire
```











