---
title: "Lab 8"
author: "Christian Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 8 Tasks

## Task 1

```{r}
setwd("C:/Users/cmill/OneDrive/Desktop/Applied Stats/Labs/Lab 8")
getwd()
```

## Task 2

```{r}
sample = runif(n = 10, min = 0, max = 5)
sample
```

Population mean and variance for uniform distribution where a = 0, b = 5 :

```{r}
mu = (0 + 5)/2
mu
```

```{r}
sigmasq = (5 - 0)^2/12
sigmasq
```

Mean and variance from sample:

```{r}
xbar = mean(sample)
xbar

s_squared = var(sample)
s_squared
```

Though some variability, the sample mean and variance are similar to the population parameters mu and sigma^2. However, with a larger sample size, the variability between the sample statistics and population parameters would lessen, according to the Central Limit Theorem.

From what we are given, we can deduce that $E(\bar{Y}) = E(Y_i) = \mu = \frac{a+b}{2}$, $E(T) = nE(Y_i) = n\mu = n\frac{a+b}{2}$, 

$$
V(\bar{Y}) = \frac{1}{n} V(Y_i) = \frac{1}{n} \sigma^2 = \frac{(b-a)^2}{12n},
$$
and 
$$
V(T) = nV(Y_i) = n\sigma^2 = n\frac{(b-a)^2}{12}.
$$





```{r}
myclt=function(n,iter){
y=runif(n*iter, 0, 5) # (A)
data=matrix(y,nr=n,nc=iter,byrow=TRUE) # (B)
sm=apply(data,2,sum) # (C)
hist(sm)
sm
}
w=myclt(n=10,iter=10000) # (D)
```

In the function defined above:

(A) creates a random sample from a normal distribution with size n*iter, a lower limit of 0, and an upper limit of 5, and stores it into an object, y.

(B) places these numbers into a matrix, where the columns will correspond to the iteration, and the rows will equal the sample size n.

(C) sums the columns (samples) of the matrix, and these are placed in a vector sm.

(D) runs the function and stores the output into a vector w, and returns a histogram of sm.

```{r}
w=myclt(n=10,iter=10000)
```

```{r}
mean(w)
```
```{r}
var(w)
```


```{r}
myclt.mean = function(n,iter){

  y = runif(n*iter, 0, 5) # (A)

  data = matrix(y,nr=n,nc=iter,byrow=TRUE) # (B)

  m = apply(data,2,mean) # (C)

  hist(m)

  m
}

w = myclt.mean(n=10,iter=10000) # (D)
```

```{r}
mean(w)
```

```{r}
var(w)
```

## Task 3

```{r}
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)

}
mycltu(n=20,iter=100000)
```

The `apply(data,2,mean)` part within `mycltu()` uses 2 for the `MARGIN` argument in `apply()` to indicate that the mean will be calculated using the columns of the matrix "data".

When applying `mycltu(n=20,iter=100000)`, there are 10,000 terms stored in the vector `w`.

`sd` takes the formula as shown in the `mycltu()` function because we know that the variance of a uniform distribution is given by $\sigma_{Y}^2 = \frac{(b-a)^2}{12}$, which tells us that $\sigma_{Y} = \frac{b-a}{\sqrt{12}}$. Because \[V(\bar{Y}) = \frac{1}{n} V(Y), \text{ we can also say that } \sigma_{\bar{Y}} = \frac{b-a}{\sqrt{12n}}.
\]

```{r}
mycltu(n = 1, iter = 10000, a = 0, b = 10)

mycltu(n = 2, iter = 10000, a = 0, b = 10)

mycltu(n = 3, iter = 10000, a = 0, b = 10)

mycltu(n = 5, iter = 10000, a = 0, b = 10)

mycltu(n = 10, iter = 10000, a = 0, b = 10)

mycltu(n = 30, iter = 10000, a = 0, b = 10)
```

The plots above demonstrate the properties of Central Limit Theorem, showing how the distribution converges to normal as the sample size increases. At n=30, the observed density, displayed conveniently with the blue curve, almost perfectly overlaps with the theoretical density, the dashed red curve.


## Task 4

```{r}
mycltb=function(n,iter,p=0.5,...){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}
```

```{r}
mycltb(n = 4, iter = 10000, p = 0.3)

mycltb(n = 5, iter = 10000, p = 0.3)

mycltb(n = 10, iter = 10000, p = 0.3)

mycltb(n = 20, iter = 10000, p = 0.3)
```

```{r}
mycltb(n = 4, iter = 10000, p = 0.7)

mycltb(n = 5, iter = 10000, p = 0.7)

mycltb(n = 10, iter = 10000, p = 0.7)

mycltb(n = 20, iter = 10000, p = 0.7)
```

```{r}
mycltb(n = 4, iter = 10000, p = 0.5)

mycltb(n = 5, iter = 10000, p = 0.5)

mycltb(n = 10, iter = 10000, p = 0.5)

mycltb(n = 20, iter = 10000, p = 0.5)
```

Like the uniform distribution case in Task 2, making samples from a binomial distribution shows the effect of the Central Limit Theorem. In each case of p=0.3, p=0.7, p=0.5, the distribution more closely resembles a normal distribution as the sample size increases. However, there is still some variability between the observed densities and the theoretical density at a sample size of n=20.


## Task 5

```{r}
mycltp=function(n,iter,lambda=10,...){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}
```

```{r}
mycltp(n = 2, iter = 10000, lambda = 4)

mycltp(n = 3, iter = 10000, lambda = 4)

mycltp(n = 5, iter = 10000, lambda = 4)

mycltp(n = 10, iter = 10000, lambda = 4)

mycltp(n = 20, iter = 10000, lambda = 4)
```

```{r}
mycltp(n = 2, iter = 10000, lambda = 10)

mycltp(n = 3, iter = 10000, lambda = 10)

mycltp(n = 5, iter = 10000, lambda = 10)

mycltp(n = 10, iter = 10000, lambda = 10)

mycltp(n = 20, iter = 10000, lambda = 10)
```


## Task 6

```{r}
MATH4753CSM24::mycltp(n = 20, iter = 10000, lambda = 4)
```




























