---
title: "Lab 6"
author: "Christian Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 6 Tasks

## Task 1
```{r}
getwd()
```

## Task 2: The Normal Distribution

### Plot Normal Curves
```{r}
x <- NULL
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dnorm(x, mean=10,sd=4),xlim=c(10 - 3*4, 10 + 3*4))

curve(dnorm(x, mean=10,sd=2),xlim=c(10 - 3*2, 10 + 3*2))

curve(dnorm(x, mean=5,sd=10),xlim=c(5 - 3*10, 5 + 3*10))

curve(dnorm(x, mean=5,sd=0.5),xlim=c(5 - 3*0.5, 5 + 3*0.5))
```

### Plot Regions and Probabilities

$$
Y \sim \mathcal{N}(0, 1), \quad P(Y \geq 2)
$$

```{r}
x <- NULL
curve(dnorm(x, 0, 1), xlim = c(0 - 3*1, 0 + 3*1))

# Define the range of the area to highlight 
xcurve <- seq(2, 10, length = 1000) # Length defines the precision (1000 is standard practice)

# Calculate the density of the normal distribution for this range
ycurve <- dnorm(xcurve, 0, 1)

# Shade the area under the curve between x = 6 and x = 8
polygon(c(2, xcurve, 10), c(0, ycurve, 0), col = "purple")

# Calculate the Area
area = 1 - pnorm(2, 0, 1)

# Round to Four Decimal Places
area = round(area, 4)

#Paste the Area into the Plot
text(2.5, dnorm(0, 0, 1) - 0.3, paste0("Area = ", area))
```

$$
Y \sim \mathcal{N}(\mu = 4, \sigma = 2), \quad P(1 \leq Y < 5)
$$

```{r}
x <- NULL
curve(dnorm(x, 4, 2), xlim = c(4 - 3*2, 4 + 3*2))

# Define the range of the area to highlight 
xcurve <- seq(1, 5, length = 1000) # Length defines the precision (1000 is standard practice)

# Calculate the density of the normal distribution for this range
ycurve <- dnorm(xcurve, 4, 2)

# Shade the area under the curve between x = 6 and x = 8
polygon(c(1, xcurve, 5), c(0, ycurve, 0), col = "orange")

# Calculate the Area
area = pnorm(5, 4, 2) - pnorm(1, 4, 2)

# Round to Four Decimal Places
area = round(area, 4)

#Paste the Area into the Plot
text(1.5, dnorm(4, 4, 2) - 0.01, paste0("Area = ", area))
```

$$
Y \sim \mathcal{N}(\mu = 10, \sigma = 4), \quad P(Y < 10)
$$

```{r}
x <- NULL
curve(dnorm(x, 10, 4), xlim = c(10 - 3*4, 10 + 3*4))

# Define the range of the area to highlight 
xcurve <- seq(-100, 10, length = 1000) # Length defines the precision (1000 is standard practice)

# Calculate the density of the normal distribution for this range
ycurve <- dnorm(xcurve, 10, 4)

# Shade the area under the curve between x = 6 and x = 8
polygon(c(-10, xcurve, 10), c(0, ycurve, 0), col = "red")

# Calculate the Area
area = pnorm(10, 10, 4) - pnorm(-10, 10, 4)

# Round to Four Decimal Places
area = round(area, 4)

#Paste the Area into the Plot
text(3, dnorm(10, 10, 4)-0.02, paste0("Area = ", area))
```

$$
Y \sim \mathcal{N}(\mu = -2, \sigma = 1/2), \quad P(-3 < Y \leq -2)
$$

```{r}
x <- NULL
curve(dnorm(x, -2, 0.5), xlim = c(-2 - 3*0.5, -2 + 3*0.5))

# Define the range of the area to highlight 
xcurve <- seq(-3, -2, length = 1000) # Length defines the precision (1000 is standard practice)

# Calculate the density of the normal distribution for this range
ycurve <- dnorm(xcurve, -2, 0.5)

# Shade the area under the curve between x = 6 and x = 8
polygon(c(-3, xcurve, -2), c(0, ycurve, 0), col = "blue")

# Calculate the Area
area = pnorm(-2, -2, 0.5) - pnorm(-3, -2, 0.5)

# Round to Four Decimal Places
area = round(area, 4)

#Paste the Area into the Plot
text(-2.75, dnorm(-2, -2, 0.5)-0.1, paste0("Area = ", area))
```

## Task 3: The Gamma Distribution

### Plot Gamma Curves

```{r}
curve(dgamma(x,shape=1,scale=1),xlim=c(0,10),ylim=c(0,1),col="red",lwd=2,
ylab="Gamma density", main = "Beta = 1")
text(1, dgamma(0.2, 1, 1), paste0("alpha=", 1))

curve(dgamma(x,shape=3,scale=1),xlim=c(0,10),ylim=c(0,1),col="blue", add = TRUE, lwd=2,
ylab="Gamma density")
text(3.25, dgamma(2.1, 3, 1)+0.05, paste0("alpha=", 3))

curve(dgamma(x,shape=5,scale=1),xlim=c(0,10),ylim=c(0,1),col="green", add = TRUE, lwd=2,
ylab="Gamma density")
text(6.5, dgamma(1.6, 1, 1), paste0("alpha=", 5))
```

### Plot Regions and Probabilities

```{r}
x <- NULL
curve(dgamma(x,shape=3,scale=2),xlim=c(0,10),ylim=c(0,1),col="black",lwd=2,
ylab="Gamma density")
text(8, dgamma(0.25, 3, 2), paste0("alpha=", 3))
# x values corresponding to the x - cords of points on the curve
xcurve=seq(2,5, length = 1000)
ycurve=dgamma(xcurve, 3, scale = 2)
polygon(c(2, xcurve, 5), c(0, ycurve, 0), col = "lightgrey")
area = pgamma(5, 3, scale = 2) - pgamma(2, 3, scale = 2)
text(3.5, 0.22, paste0("Area =", round(area, 4)))
```

```{r}
x <- NULL
curve(dgamma(x,shape=6,scale=3),xlim=c(0,10),ylim=c(0,1),col="black",lwd=2,
ylab="Gamma density")
text(8, dgamma(0.7, 6, 3), paste0("alpha=", 6))
# x values corresponding to the x - cords of points on the curve
xcurve=seq(1,4, length = 1000)
ycurve=dgamma(xcurve, 6, scale = 3)
polygon(c(1, xcurve, 4), c(0, ycurve, 0), col = "lightgrey")
area = pgamma(4, 6, scale = 3) - pgamma(1, 6, scale = 3)
text(3.5, 0.08, paste0("Area =", round(area, 4)))
```

```{r}
x <- NULL
curve(dgamma(x,shape=2,scale=4),xlim=c(0,10),ylim=c(0,1),col="black",lwd=2,
ylab="Gamma density")
text(8, dgamma(6, 2, 4)+0.15, paste0("alpha=", 2))
# x values corresponding to the x - cords of points on the curve
xcurve=seq(3,6, length = 1000)
ycurve=dgamma(xcurve, 2, scale = 4)
polygon(c(3, xcurve, 6), c(0, ycurve, 0), col = "lightgrey")
area = pgamma(6, 2, scale = 4) - pgamma(3, 2, scale = 4)
text(4.5, 0.18, paste0("Area =", round(area, 4)))
```

##  Task 4: The Chi-Square Distribution

### Plot Chi-Square Curves
```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dchisq(x,df=1),xlim=c(0,10),ylim=c(0,1),col="red",lwd=2,
ylab="Chisq density", main="df=1")

curve(dchisq(x,df=2),xlim=c(0,10),ylim=c(0,1),col="blue",lwd=2,
ylab="Chisq density", main = "df = 2")

curve(dchisq(x,df=4),xlim=c(0,10),ylim=c(0,1),col="purple",lwd=2,
ylab="Chisq density", main = "df = 4")

curve(dchisq(x,df=20),xlim=c(0,25),ylim=c(0,1),col="green",lwd=2,
ylab="Chisq density", main = "df = 20")

```

### Plot Regions and Probabilities

```{r}
curve(dchisq(x,df=2),xlim=c(0,10),ylim=c(0,1),col="blue",lwd=2,
ylab="Chisq density", main = "df = 2")

xcurve=seq(2,4, length = 1000)
ycurve=dchisq(xcurve, df = 2)
polygon(c(2, xcurve, 4), c(0, ycurve, 0), col = "lightgrey")
area = pchisq(4, df = 2) - pchisq(2, df = 2)
text(3.2, 0.21, paste0("Area =", round(area, 4)))
```


```{r}
curve(dchisq(x,df=3),xlim=c(0,10),ylim=c(0,1),col="orange",lwd=2,
ylab="Chisq density", main = "df = 3")

xcurve=seq(3,5, length = 1000)
ycurve=dchisq(xcurve, df = 3)
polygon(c(3, xcurve, 5), c(0, ycurve, 0), col = "lightgrey")
area = pchisq(5, df = 3) - pchisq(3, df = 3)
text(4, 0.2, paste0("Area =", round(area, 4)))

```

```{r}
curve(dchisq(x,df=20),xlim=c(0,25),ylim=c(0,1),col="green",lwd=2,
ylab="Chisq density", main = "df = 20")

xcurve=seq(10,21, length = 1000)
ycurve=dchisq(xcurve, df = 20)
polygon(c(10, xcurve, 21), c(0, ycurve, 0), col = "lightgrey")
area = pchisq(21, df = 20) - pchisq(10, df = 20)
text(16, 0.15, paste0("Area =", round(area, 4)))
```

## Task 5: The Weibull Distribution

### Plot Weibull Curves

```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dweibull(x,shape=2,scale=1),xlim=c(0,2.5),ylim=c(0,2),col="red",lwd=2,
ylab="Weibull Density", main= "Shape = 2, Scale = 1")

curve(dweibull(x,shape=4,scale=1),xlim=c(0,2.5),ylim=c(0,2),col="blue",lwd=2,
ylab="Weibull Density", main = "Shape = 4, Scale = 1")

curve(dweibull(x,shape=2,scale=2),xlim=c(0,2.5),ylim=c(0,2),col="purple",lwd=2,
ylab="Weibull Density", main = "Shape = 2, Scale = 2")

curve(dweibull(x,shape=2,scale=4),xlim=c(0,.25),ylim=c(0,2),col="green",lwd=2,
ylab="Weibull Density", main = "Shape = 2, Scale = 4")
```

### Plot Regions and Probabilities

```{r}
curve(dweibull(x,shape=2,scale=1),xlim=c(0,2.5),ylim=c(0,2),col="red",lwd=2,
ylab="Weibull Density", main= "Shape = 2, Scale = 1")

xcurve=seq(0.5,1, length = 1000)
ycurve=dweibull(xcurve, 2, scale = 1)
polygon(c(0.5, xcurve, 1), c(0, ycurve, 0), col = "lightgrey")
area = pweibull(1, 2, scale = 1) - pweibull(0.5, 2, scale = 1)
text(0.75, 1, paste0("Area =", round(area, 4)))
```

```{r}
curve(dweibull(x,shape=4,scale=1),xlim=c(0,2.5),ylim=c(0,2),col="blue",lwd=2,
ylab="Weibull Density", main= "Shape = 4, Scale = 1")

xcurve=seq(1, 1.5, length = 1000)
ycurve=dweibull(xcurve, 4, scale = 1)
polygon(c(1, xcurve, 1.5), c(0, ycurve, 0), col = "lightgrey")
area = pweibull(1.5, 4, scale = 1) - pweibull(1, 4, scale = 1)
text(1.5, 1, paste0("Area =", round(area, 4)))
```

```{r}
curve(dweibull(x,shape=2,scale=2),xlim=c(0,2.5),ylim=c(0,2),col="purple",lwd=2,
ylab="Weibull Density", main= "Shape = 2, Scale = 2")

xcurve=seq(0.5, 2, length = 1000)
ycurve=dweibull(xcurve, 2, scale = 2)
polygon(c(0.5, xcurve, 2), c(0, ycurve, 0), col = "lightgrey")
area = pweibull(2, 2, scale = 2) - pweibull(0.5, 2, scale = 2)
text(1.25, 0.6, paste0("Area =", round(area, 4)))
```

## Task 6: The Beta Distribution

### Plot Beta Curves

```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dbeta(x,shape1=2,shape2=4),xlim=c(0,1),ylim=c(0,2.2),col="red",lwd=2,
ylab="Beta Density", main= "Shape1 = 2, Shape2 = 4")

curve(dbeta(x,shape1=3,shape2=2),xlim=c(0,1),ylim=c(0,2.2),col="blue",lwd=2,
ylab="Beta Density", main = "Shape1 = 3, Shape2 = 2")

curve(dbeta(x,shape1=2,shape2=2),xlim=c(0,1),ylim=c(0,2.2),col="purple",lwd=2,
ylab="Beta Density", main = "Shape1 = 2, Shape2 = 2")

curve(dbeta(x,shape1=2,shape2=3),xlim=c(0,1),ylim=c(0,2.2),col="green",lwd=2,
ylab="Beta Density", main = "Shape1 = 2, Shape2 = 3")
```

### Plot Regions and Probabilities

```{r}
curve(dbeta(x,shape1=2,shape2=4),xlim=c(0,1),ylim=c(0,2.2),col="red",lwd=2,
ylab="Beta Density", main= "Shape1 = 2, Shape2 = 4")

xcurve=seq(0.5, 2, length = 1000)
ycurve=dbeta(xcurve, shape1=2,shape2=4)
polygon(c(0.5, xcurve, 2), c(0, ycurve, 0), col = "lightgrey")
area = pbeta(2,shape1=2,shape2=4) - pbeta(0.5,shape1=2,shape2=4)
text(0.75, 0.7, paste0("Area =", round(area, 4)))
```

```{r}
curve(dbeta(x,shape1=3,shape2=2),xlim=c(0,1),ylim=c(0,2.2),col="blue",lwd=2,
ylab="Beta Density", main = "Shape1 = 3, Shape2 = 2")

xcurve=seq(0.2, 1, length = 1000)
ycurve=dbeta(xcurve,shape1=3,shape2=2)
polygon(c(0.2, xcurve, 1), c(0, ycurve, 0), col = "lightgrey")
area = pbeta(1,shape1=3,shape2=2) - pbeta(0.2,shape1=3,shape2=2)
text(0.7, 1.92, paste0("Area =", round(area, 4)))
```

```{r}
curve(dbeta(x,shape1=2,shape2=3),xlim=c(0,1),ylim=c(0,2.2),col="green",lwd=2,
ylab="Beta Density", main = "Shape1 = 2, Shape2 = 3")

xcurve=seq(0.2, 0.4, length = 1000)
ycurve=dbeta(xcurve,shape1=2,shape2=3)
polygon(c(0.2, xcurve, 0.4), c(0, ycurve, 0), col = "lightgrey")
area = pbeta(0.4,shape1=2,shape2=3) - pbeta(0.2,shape1=2,shape2=3)
text(0.3, 1.95, paste0("Area =", round(area, 4)))
```

## Task 7: myncurve() Function

```{r}
MATH4753CSM24::myncurve(mu = 10, sigma = 5, a = 6, plot = TRUE)
```




